Noun Serialization

### 2025-12-01 mss.tex
Fix +gel.

@@

\begin{table}[ht]
\centering
\caption{Example noun encodings (\texttt{+gel} encoding).  \texttt{.} dot delimits bit fields for readability.}
\label{tab:gel-encodings}
\begin{tabular}{ r r }
  \multicolumn{1}{l}{\textbf{Noun}} & \multicolumn{1}{l}{\textbf{Encoded Form}} \\
  \hline
  \texttt{[0 0]} & \texttt{0b1.0.1.0.1} \\
  \texttt{[1 0]} & \texttt{0b110.0.1.0.1} \\
  \texttt{[2 1 0]} & \texttt{0b110.0.1.0.1.100100.0.1} \\
  \texttt{[[2 3] 1 0]} & \texttt{0b110.0.1.0.1.110100.0.100100.0.1.1} \\
\end{tabular}
\end{table}

Our code implementation for \texttt{+gel} is as follows:

\begin{lstlisting}[style=listingcode]
!:  |%
++  gel
  =gel !:  |=  a=*
  ^-  @
  =+  l=0
  =+  b=0
  =<  -
  |-
  ?^  a
    =+  lv=$(a -.a)
    =+  rv=$(a +.a)
    =+  [c l]=(mash rv lv)
    [(con (lsh [0 1] c) 0b1) +(l)]
  ?:  =(0 a)  [0b10 4]
  :: need another mash in here for unary length
  =+  [c l]=(mash [a l] [b +((met 0 b))])
  [(con (lsh [0 1] c) 0b0) +(l)]
:: length of atom in unary
++  len
  |=  a=@
  ^-  @
  (fil 0 (met 0 a) 0b1)
:: mash two atoms together
++  mash
  |=  [a=[p=@ l=@] b=[p=@ l=@]]
  ^-  [c=@ l=@]
  :-  (con (lsh [0 l.b] p.a) p.b)
  (add l.a l.b)
--
\end{lstlisting}

@@

\begin{lstlisting}[style=listingcode]
++  jam
  ~/  %jam
  |=  a=*
  ^-  @
  =+  b=0
  =+  m=`(map * @)`~
  =<  q
  |-  ^-  [p=@ q=@ r=(map * @)]
  =+  c=(~(get by m) a)
  ?~  c
    =>  .(m (~(put by m) a b))
    ?:  ?=(@ a)
      =+  d=(mat a)
      [(add 1 p.d) (lsh 0 q.d) m]
    =>  .(b (add 2 b))
    =+  d=$(a -.a)
    =+  e=$(a +.a, b (add b p.d), m r.d)
    :+  (add 2 (add p.d p.e))
      (mix 1 (lsh [0 2] (cat 0 q.d q.e)))
    r.e
  ?:  ?&(?=(@ a) (lte (met 0 a) (met 0 u.c)))
    =+  d=(mat a)
    [(add 1 p.d) (lsh 0 q.d) m]
  =+  d=(mat u.c)
  [(add 2 p.d) (mix 3 (lsh [0 2] q.d)) m]
::  length encode
++  mat
  ~/  %mat
  |=  a=@
  ^-  [p=@ q=@]
  ?:  =(0 a)
    [1 1]
  =+  b=(met 0 a)
  =+  c=(met 0 b)
  :-  (add (add c c) b)
  (cat 0 (bex c) (mix (end [0 (dec c)] b) (lsh [0 (dec c)] a)))
\end{lstlisting}

>>>>

The reference implementation of \texttt{+mat} encoding in Hoon is as follows:

\begin{lstlisting}[style=listingcode]
++  mat
  ~/  %mat
  |=  a=@
  ^-  [p=@ q=@]
  ?:  =(0 a)
    [1 1]
  =+  b=(met 0 a)
  =+  c=(met 0 b)
  :-  (add (add c c) b)
  (cat 0 (bex c) (mix (end [0 (dec c)] b)
   ↳                  (lsh [0 (dec c)] a)))
\end{lstlisting}

@@

\begin{table}[ht]
\centering
\caption{Example noun encodings (\texttt{+gel} encoding).  \texttt{.} dot delimits bit fields for readability.}
\label{tab:gel-encodings}
\begin{tabular}{ r r }
  \multicolumn{1}{l}{\textbf{Noun}} & \multicolumn{1}{l}{\textbf{Encoded Form}} \\
  \hline
  \texttt{0} & \texttt{0b0.1.0} \\
  \texttt{[0 0]} & \texttt{0b1.0.1.0.1} \\
  \texttt{[1 0]} & \texttt{0b1.0.110.0.1} \\
  \texttt{[2 1 0]} & \texttt{0b1.0.110.0.1.100100.0.1} \\
  \texttt{[[2 3] 1 0]} & \texttt{0b1.0.110.0.1.110100.0.100100.0.1.1} \\
\end{tabular}
\end{table}

\newpage
Our code implementation for \texttt{+gel} encoding is as follows:

\begin{lstlisting}[style=listingcode]
++  gel
  |=  a=*
  ^-  @
  =+  b=0
  =<  q
  |-  ^-  [p=@ q=@]
  ?:  ?=(@ a)
    =+  d=(mat a)
    [(add 1 p.d) (lsh 0 q.d)]
  =>  .(b (add 1 b))
  =+  d=$(a -.a)
  =+  e=$(a +.a, b (add b p.d))
  :-  (add 1 (add p.d p.e))
  (mix 1 (lsh [0 1] (cat 0 q.d q.e)))
\end{lstlisting}

\noindent
Decoding is left as an exercise to the interested reader.

@@

\begin{lstlisting}[style=listingcode]
++  jam
  ~/  %jam
  |=  a=*
  ^-  @
  =+  b=0
  =+  m=`(map * @)`~
  =<  q
  |-  ^-  [p=@ q=@ r=(map * @)]
  =+  c=(~(get by m) a)
  ?~  c
    =>  .(m (~(put by m) a b))
    ?:  ?=(@ a)
      =+  d=(mat a)
      [(add 1 p.d) (lsh 0 q.d) m]
    =>  .(b (add 2 b))
    =+  d=$(a -.a)
    =+  e=$(a +.a, b (add b p.d), m r.d)
    :+  (add 2 (add p.d p.e))
      (mix 1 (lsh [0 2] (cat 0 q.d q.e)))
    r.e
  ?:  ?&(?=(@ a) (lte (met 0 a) (met 0 u.c)))
    =+  d=(mat a)
    [(add 1 p.d) (lsh 0 q.d) m]
  =+  d=(mat u.c)
  [(add 2 p.d) (mix 3 (lsh [0 2] q.d)) m]
\end{lstlisting}
